Evaluating approaches to deriving photometric redshift posterior probability distribution functions for \textsc{LSST}

### 2019-03-10 paper/tex/methods.tex
more concise CMNN discussion

The nearest-neighbours colour-matching photometric redshift estimator (\textsc{CMNN}) is presented in \citet[][hereafter G18]{Graham:17}.
This method uses a training set of galaxies with known redshifts that has equivalent or better photometry as the test set in terms of quality and filter coverage.
For each galaxy in the test set we identify a colour-matched subset of training galaxies.
This subset is identified by first calculating the Mahalanobis distance $D_M$ in colour-space between the test galaxy and all training-set galaxies:
\begin{equation}
D_{\rm M} = \sum^{N_{\rm colours}}\frac{(c_{\rm train}-c_{\rm test})^2}{(\delta c_{\rm test})^2}
\end{equation}
where c colour, $\delta c_{\rm test}$ is the measurement error on the colour, and $N_{\rm colours}$ is the total number of colors (i.e., in our case $u$-$g$, $g$-$r$, $r$-$i$, $i$-$z$, and $z$-$y$).
Then, we choose a threshold value for $D_M$ that defines the colour match set based on a set value of the percent point function (PPF): for example, for $N_{\rm dof}=5$, choosing a vPPF$=0.95$, $95$ per cent of all training galaxies consistent with the test galaxy will have $D_M < 11.07$ (where $N_{\rm dof}$ is the number of degrees of freedom, in this case the number of colours).
If a galaxy had a non-detection in a band, that band was dropped and $N_{\rm dof}$ was reduced by one in the colour-matching space.
For a given test galaxy, the $p(z)$ is the normalized distribution of the true catalogue redshifts of this colour-matched subset of training galaxies.

We have applied the nearest-neighbours colour-matching photometric redshift estimator described in G18 to the simulated data.
Compared to its application in G18, there are some minor differences in the application of this estimator to the Buzzard catalogue.
First, we do not impose non-detections on galaxies with a magnitude fainter than the expected LSST 10-year limiting magnitude or bright enough to saturate with LSST: {\it all} of the photometry for all the galaxies in the test and training sets are used for this experiment.
Second, as in G18 we do apply an initial cut in colour to the training set before calculating the Mahalanobis distance in order to accelerate processing, and also use a magnitude pseudo-prior to improve photo-$z$ estimates, but for both we have used different cut-off values that are appropriate for the Buzzard galaxies' colours and magnitudes.
Third, we set different parameters for the identification of the colour-matched subset of training galaxies and the selection of a photometric redshift estimate.
In G18 we used a PPF value of 0.68 to identify the colour-matched subset of training galaxies and used the redshift of nearest neighbour in colour-space as the photo-$z$ estimate.
These choices work well when the desire is to obtain accurate photo-$z$ estimates for most test-set galaxies, but does not return an accurate $p(z)$ in all cases -- especially for galaxies that are bright and/or have few matches in colour-space.
Since an accurate estimate of $p(z)$ is desired for this work we make several changes to our implementation of the \textsc{CMNN} photo-$z$ estimator.
We continue to use a ${\rm PPF} = 0.95$ to generate the subset of colour-matched training galaxies, but weight them by the inverse of their Mahalanobis distance.
This weighting maintains some of the accuracy that was previously achieved by simply using the nearest neighbour in colour-space.
We then use the weights to create the $p(z)$ instead of having the redshift of each colour-matched training-set galaxy count equally.
To obtain a robust estimate of the $p(z)$ for galaxies with a small number of colour-matched training set galaxies, when this number is less than $20$ the nearest $20$ neighbours in colour-space are used instead, and we convolve the $p(z)$ with a Gaussian with a standard deviation of:
\begin{equation}
\sigma = \sigma_{\rm train} \sqrt{({\rm PPF}_{20}/0.95)^2 -1}
\end{equation}
\noindent to appropriately broaden it so that the $p(z)$ for these test galaxies represents the enlarged PPF value associated with it.
Overall, these three changes will yield less precise photo-$z$ estimates compared to those presented in G18, but they will all have significantly more accurate estimates of the $p(z)$, particularly for the brightest test galaxies.
This is sufficient for this work because, as described in G18, the goal of the \textsc{CMNN} photo-$z$ estimator was never to provide the ``best'' (or even competitive) estimates in the first place, given its reliance on a deep training set, but rather to provide a means for direct comparisons between LSST photometric quality and photo-$z$ estimates.
With this work we show how the input parameters should be set in order to return accurate $p(z)$ estimates in addition to point value estimates.

>>>>

The nearest-neighbours colour-matching photometric redshift estimator \citep[\cmnn][]{Graham:17} uses a training set of galaxies with known redshifts that has equivalent or better photometry than the test set in terms of quality and filter coverage.
For each galaxy in the test set, \cmnn\ identifies a colour-matched subset of training galaxies using a threshold in the Mahalanobis distance $D_M$ in the space of available colours for the test set galaxy
% between the test galaxy and all training set galaxies:
% \begin{equation}
% D_{\rm M} = \sum^{N_{\rm colours}}\frac{(c_{\rm train}-c_{\rm test})^2}{(\delta c_{\rm test})^2}
% \end{equation}
% where c colour, $\delta c_{\rm test}$ is the measurement error on the colour, and $N_{\rm colours}$ is the total number of colors (i.e., in our case $u$-$g$, $g$-$r$, $r$-$i$, $i$-$z$, and $z$-$y$).
that defines the colour match set based on a value of the percent point function (PPF).
As an example, for $N_{\rm{filt}}=5$ with PPF$=0.95$, $95\%$ of all training galaxies consistent with the test galaxy will have $D_M < 11.07$.
Undetected bands are dropped, thereby reducing the effective $N_{\rm{filt}}$ for that galaxy.
The \pzpdf\ of a given test set galaxy is the normalized distribution of redshifts of its colour-matched subset of training set galaxies.

Here, we make two modifications to the implementation of \citet{Graham:17} to comply with the controlled experimental conditions.
First, we do not impose non-detections on galaxies fainter than the expected \lsst\ 10-year limiting magnitude or bright enough to saturate with \lsst's CCDs, instead using all of the photometry for the DC1 test and training sets.
Second, we apply the initial colour cut to the training set before calculating the Mahalanobis distance in order to accelerate processing and use a magnitude pseudo-prior as in \citet{Graham:17}, but for both we use cut-off values corresponding to the DC1 training set galaxies' colours and magnitudes.

We make an additional adaptation to enable the \cmnn\ algorithm to yield accurate \pzpdf s for all galaxies, as the original \citet{Graham:17} algorithm is optimized for \pz\ point estimates and is susceptible to less accurate \pzpdf s for bright galaxies or those with few matches in colour-space.
We use PPF$=0.95$ rather than PPF$=0.68$ to generate the subset of colour-matched training galaxies, whose redshifts are weighted by their inverse Mahalanobis distances of the when composing the \pzpdf\ rather than weighting all colour-matched training galaxies equally.
Additionally, when the number of colour-matched training set galaxies is less than 20, the nearest 20 neighbours in color-space are used instead, and the output \pzpdf\ is convolved with a Gaussian kernel of variance $\sigma_{\rm train}^{2}({\rm PPF}_{20}/0.95)^2 -1$ to account foe the corresponding growth of the effective PPF to include 20 neighbors.
